---
title: "Housing Cri.. Prices"
subtitle: "Data Exploration"
author: "Marc Schwering"
output:
  html_document:
    number_sections: true
    toc: true
    toc_depth: 2
    fig_width: 7
    fig_height: 4.5
    theme: cosmo
---







# Introduction



**Loading Data**

The following libraries will be used.

```{r, message=FALSE, warning=FALSE, results='hide'}
library(data.table) # database-like tables
library(ggplot2) # general plots
library(corrplot) # corr SPLOMs
library(vcd) # stats for categories
library(mice) # for imputation
```

Input is loaded and covnerted to `data.table`.

```{r}
test <- read.csv("../input/test.csv")
train <- read.csv("../input/train.csv")
test <- data.table(test)
train <- data.table(train)
```






***







# Data Quality

The number of dimensions and density of the input matrix is calculated.

```{r}
(N <- nrow(train))
(p <- ncol(train) - 2)
N^(1/p)
```

## Variable Types

From the *data_description* variable types can be defined.
There are 4 categories: *nominal*, *ordinal*, *discrete*, *continuous*.

```{r}
nomVars <- c("MSZoning", "Street", "Alley", "LandContour", "LotConfig",
             "Neighborhood", "Condition1", "Condition2", "HouseStyle",
             "RoofStyle", "RoofMatl", "Exterior1st", "Exterior2nd",
             "MasVnrType", "Foundation", "Heating", "CentralAir", "Electrical",
             "GarageType", "PavedDrive", "MiscFeature", "SaleType",
             "SaleCondition")
ordVars <- c("MSSubClass", "LotShape", "Utilities", "LandSlope", "BldgType",
             "OverallQual", "OverallCond", "ExterQual", "ExterCond", "BsmtQual",
             "BsmtCond", "BsmtExposure", "BsmtFinType1", "BsmtFinType2",
             "HeatingQC", "KitchenQual", "Functional", "FireplaceQu",
             "GarageFinish", "GarageQual", "GarageCond", "PoolQC", "Fence",
             "MoSold")
discVars <- c("YearBuilt", "YearRemodAdd", "BsmtFullBath", "BsmtHalfBath",
              "FullBath", "HalfBath", "BedroomAbvGr", "KitchenAbvGr",
              "TotRmsAbvGrd", "Fireplaces", "GarageYrBlt", "GarageCars",
              "YrSold")
contVars <- c("LotArea", "MasVnrArea", "BsmtFinSF1", "BsmtFinSF2", "BsmtUnfSF",
              "TotalBsmtSF", "X1stFlrSF", "X2ndFlrSF", "LowQualFinSF",
              "GrLivArea", "GarageArea", "WoodDeckSF", "OpenPorchSF",
              "EnclosedPorch", "X3SsnPorch", "ScreenPorch", "PoolArea",
              "MiscVal", "LotFrontage")
```

## Special Values

According to *data_description* several $NA$ have a special meaning.
These values are replaced with more descriptive `chr` strings.

```{r, results='hide', message=FALSE}
train[is.na(Alley), Alley := "noAccess"]
test[is.na(Alley), Alley := "noAccess"]
train[is.na(BsmtQual), BsmtQual := "noBasement"]
train[is.na(BsmtCond), BsmtCond := "noBasement"]
train[is.na(BsmtExposure), BsmtExposure := "noBasement"]
train[is.na(BsmtFinType1), BsmtFinType1 := "noBasement"]
train[is.na(BsmtFinType2), BsmtFinType2 := "noBasement"]
test[is.na(BsmtQual), BsmtQual := "noBasement"]
test[is.na(BsmtCond), BsmtCond := "noBasement"]
test[is.na(BsmtExposure), BsmtExposure := "noBasement"]
test[is.na(BsmtFinType1), BsmtFinType1 := "noBasement"]
test[is.na(BsmtFinType2), BsmtFinType2 := "noBasement"]
train[is.na(FireplaceQu), FireplaceQu := "noFireplace"]
test[is.na(FireplaceQu), FireplaceQu := "noFireplace"]
train[is.na(GarageType), GarageType := "noGarage"]
train[is.na(GarageFinish), GarageFinish := "noGarage"]
train[is.na(GarageQual), GarageQual := "noGarage"]
train[is.na(GarageCond), GarageCond := "noGarage"]
test[is.na(GarageType), GarageType := "noGarage"]
test[is.na(GarageFinish), GarageFinish := "noGarage"]
test[is.na(GarageQual), GarageQual := "noGarage"]
test[is.na(GarageCond), GarageCond := "noGarage"]
train[is.na(PoolQC) , PoolQC := "noPool"]
test[is.na(PoolQC) , PoolQC := "noPool"]
train[is.na(Fence), Fence := "noFence"]
test[is.na(Fence), Fence := "noFence"]
train[is.na(MiscFeature), MiscFeature := "none"]
test[is.na(MiscFeature), MiscFeature := "none"]
```

Nominal and ordinal variables are converted to `factor`,
continuous and discrete to `numeric` and `integer` respectively.

```{r}
for (j in colnames(train)) {
  if (j %in% nomVars || j %in% ordVars) {
    train[[j]] <- as.factor(train[[j]])
    test[[j]] <- as.factor(test[[j]])
  } else if (j %in% discVars) {
    train[[j]] <- as.integer(train[[j]])
    test[[j]] <- as.integer(test[[j]])
  } else if (j %in% contVars) {
    train[[j]] <- as.numeric(train[[j]])
    test[[j]] <- as.numeric(test[[j]])
  }
}
```

## Ordering

From *data_description* the order of ordinal variable classes can be derived.
These orderings are defined, then the `levels` of their `factors` are adjusted.

```{r}
lvlOrd <- list(
  LotShape = c("Reg", "IR1", "IR2", "IR3"),
  Utilities = c("AllPub", "NoSewr", "NoSeWa", "ELO"),
  LandSlope = c("Gtl", "Mod", "Sev"),
  BldgType = c("1Fam", "2FmCon", "Duplx", "TwnhsE", "TwnhsI"),
  OverallQual = 10:1,
  OverallCond = 10:1,
  ExterQual = c("Ex", "Gd", "TA", "Fa", "Po"),
  ExterCond = c("Ex", "Gd", "TA", "Fa", "Po"),
  BsmtQual = c("Ex", "Gd", "TA", "Fa", "Po", "noBasement"),
  BsmtCond = c("Ex", "Gd", "TA", "Fa", "Po", "noBasement"),
  BsmtExposure = c("Gd", "Av", "Mn", "No", "noBasement"),
  BsmtFinType1 = c("GLQ", "ALQ", "BLQ", "Rec", "LwQ", "Unf", "noBasement"),
  BsmtFinType2 = c("GLQ", "ALQ", "BLQ", "Rec", "LwQ", "Unf", "noBasement"),
  HeatingQC = c("Ex", "Gd", "TA", "Fa", "Po"),
  KitchenQual = c("Ex", "Gd", "TA", "Fa", "Po"),
  Functional = c("Typ", "Min1", "Min2", "Mod", "Maj1", "Maj2", "Sev", "Sal"),
  FireplaceQu = c("Ex", "Gd", "TA", "Fa", "Po", "noFireplace"),
  GarageFinish = c("Fin", "RFn", "Unf", "noGarage"),
  GarageQual = c("Ex", "Gd", "TA", "Fa", "Po", "noGarage"),
  GarageCond = c("Ex", "Gd", "TA", "Fa", "Po", "noGarage"),
  PoolQC = c("Ex", "Gd", "TA", "Fa", "noPool"),
  Fence = c("GdPrv", "MnPrv", "GdWo", "MnWw", "noFence"),
  MoSold = 1:12
)
for (j in names(lvlOrd)) {
  train[[j]] <- factor(train[[j]], levels = lvlOrd[[j]])
  test[[j]] <- factor(test[[j]], levels = lvlOrd[[j]])
}
```

## Missing Data

$NA$ which are left must be true missing data.
The following table gives a complete summary of missing data.

```{r}
full <- rbind(train[, -ncol(train), with = FALSE], test)
full$label <- rep(c("train", "test"), c(nrow(test), nrow(train)))
naFeats <- vapply(
  colnames(full),
  function(j) any(is.na(full[[j]])) || any(full[[j]] == ""),
  logical(1)
)
md.pattern(full[, colnames(full)[naFeats], with = FALSE])
```











***











# Feature Exploration


## Continuous and Discrete Features

### Correlatio between Features

The Pearson Correlation Coefficient is calculated for all feature pairs
where both features are discrete or continous.
The results are shown as heatmap below.
Rows and columns were hierarchically ordered (complete link).

```{r, fig.width=10, fig.height=10}
m <- cor(
  train[, c(contVars, discVars), with = FALSE],
  use = "complete.obs"
)
corrplot(m, method = "color", order = "hclust")
```

### Correlation to Target

For all continuous and discrete features the pearson correlation coefficient
to the target *SalePrice* is calculated.
Results are ordered by correlation.
The color labels the feature type, continuous or discrete.

```{r, fig.width=10, fig.height=5}
cors <- vapply(
  c(contVars, discVars),
  function(i) cor(train[[i]], train$SalePrice, use = "complete.obs"),
  numeric(1)
)
df <- data.frame(
  var = factor(c(contVars, discVars),
               levels = c(contVars, discVars)[order(cors)]),
  cor = cors,
  type = rep(c("cont", "disc"), c(length(contVars), length(discVars)))
)
ggplot(df, aes(x = var, y = cor, fill = type)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("P Correl Coeff of Cont/Discr Features vs SalePrice")
```








## Categorical Features

### Correlation between Features

The Cramer's V is calculated for all feature pairs
where both features are categories.
The results are shown as heatmap below.
Rows and columns were hierarchically ordered (complete link).

```{r, fig.width=10, fig.height=10, warning=FALSE}
m <- as.matrix(train[, c(ordVars, nomVars), with = FALSE])
m <- m[complete.cases(m), ]
mCramer <- vapply(
  seq_len(ncol(m)),
  function(i) {
    vapply(
      seq_len(ncol(m)),
      function(j) {
        assocstats(table(m[, i], m[, j]))$cramer
      },
      numeric(1)
    )
  },
  numeric(ncol(m))
)
colnames(mCramer) <- colnames(m)
rownames(mCramer) <- colnames(m)
corrplot(mCramer, method = "color", order = "hclust")
```

### Correlation to Target

For all categorical features the Kruskal-Wallis $\chi^2$ statistic
to the target *SalePrice* is calculated.
Results are ordered by the $\chi^2$ value.
The color labels the feature type, nominal or ordinal.

```{r, fig.width=10, fig.height=5, warning=FALSE}
kwX2 <- vapply(
  c(ordVars, nomVars),
  function(j) { kruskal.test(train$SalePrice, train[[j]])$statistic },
  numeric(1)
)
df <- data.frame(
  var = factor(c(ordVars, nomVars),
               levels = c(ordVars, nomVars)[order(kwX2)]),
  kwX2 = kwX2,
  type = rep(c("ord", "nom"), c(length(ordVars), length(nomVars)))
)
ggplot(df, aes(x = var, y = kwX2, fill = type)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("KW-X^2 for Categories vs SalePrice")
```







***

# Summary
