---
title: "Housing Cri.. Prices"
subtitle: "Data Exploration"
author: "Marc Schwering"
output:
  html_document:
    number_sections: true
    toc: true
    toc_depth: 2
    fig_width: 10
    fig_height: 7
    theme: cosmo
---







# Introduction



**Loading Data**

The following libraries will be used.

```{r, message=FALSE, warning=FALSE, results='hide'}
library(data.table) # database-like tables
library(ggplot2) # general plots
library(highcharter) # interactive plots
library(DT) # display tables
library(corrplot) # corr SPLOMs
library(vcd) # stats for categories
library(mice) # for imputation
```

Input is loaded and covnerted to `data.table`.

```{r}
test <- read.csv("../input/test.csv")
train <- read.csv("../input/train.csv")
test <- data.table(test)
train <- data.table(train)
```








***







# Data Quality

The number of dimensions and sampling density is calculated.
The datasets are combined.

```{r}
(N <- nrow(train))
(p <- ncol(train) - 2)
N^(1/p)
test$SalePrice <- NaN
full <- rbind(train, test)
full$label <- rep(c("train", "test"), c(nrow(train), nrow(test)))
```

## Feature Attributes

From the *data_description* variable types can be defined.
There are 4 categories: *nominal*, *ordinal*, *discrete*, *continuous*.
A lot of the discrete variables could arguably be handled as ordinals.

```{r}
nomVars <- c("MSZoning", "Street", "Alley", "LandContour", "LotConfig",
             "Neighborhood", "Condition1", "Condition2", "HouseStyle",
             "RoofStyle", "RoofMatl", "Exterior1st", "Exterior2nd",
             "MasVnrType", "Foundation", "Heating", "CentralAir", "Electrical",
             "GarageType", "PavedDrive", "MiscFeature", "SaleType",
             "SaleCondition")
ordVars <- c("MSSubClass", "LotShape", "Utilities", "LandSlope", "BldgType",
             "OverallQual", "OverallCond", "ExterQual", "ExterCond", "BsmtQual",
             "BsmtCond", "BsmtExposure", "BsmtFinType1", "BsmtFinType2",
             "HeatingQC", "KitchenQual", "Functional", "FireplaceQu",
             "GarageFinish", "GarageQual", "GarageCond", "PoolQC", "Fence",
             "MoSold")
discVars <- c("YearBuilt", "YearRemodAdd", "BsmtFullBath", "BsmtHalfBath",
              "FullBath", "HalfBath", "BedroomAbvGr", "KitchenAbvGr",
              "TotRmsAbvGrd", "Fireplaces", "GarageYrBlt", "GarageCars",
              "YrSold")
contVars <- c("LotArea", "MasVnrArea", "BsmtFinSF1", "BsmtFinSF2", "BsmtUnfSF",
              "TotalBsmtSF", "X1stFlrSF", "X2ndFlrSF", "LowQualFinSF",
              "GrLivArea", "GarageArea", "WoodDeckSF", "OpenPorchSF",
              "EnclosedPorch", "X3SsnPorch", "ScreenPorch", "PoolArea",
              "MiscVal", "LotFrontage")
```

A table is created to store feature attributes.

```{r, results='hide'}
featAttr <- data.table(
  name = colnames(test)[-1],
  type = "nominal"
)
featAttr[name %in% ordVars, type := "ordinal"]
featAttr[name %in% discVars, type := "discrete"]
featAttr[name %in% contVars, type := "continuous"]
```

**Special Values**

According to *data_description* several $NA$ have a special meaning.
These values are replaced with more descriptive `chr` strings.
Before, test and train set are combined.

```{r, results='hide', message=FALSE}
full[is.na(Alley), Alley := "noAccess"]
full[is.na(BsmtQual), BsmtQual := "noBasement"]
full[is.na(BsmtCond), BsmtCond := "noBasement"]
full[is.na(BsmtExposure), BsmtExposure := "noBasement"]
full[is.na(BsmtFinType1), BsmtFinType1 := "noBasement"]
full[is.na(BsmtFinType2), BsmtFinType2 := "noBasement"]
full[is.na(FireplaceQu), FireplaceQu := "noFireplace"]
full[is.na(GarageType), GarageType := "noGarage"]
full[is.na(GarageFinish), GarageFinish := "noGarage"]
full[is.na(GarageQual), GarageQual := "noGarage"]
full[is.na(GarageCond), GarageCond := "noGarage"]
full[is.na(PoolQC) , PoolQC := "noPool"]
full[is.na(Fence), Fence := "noFence"]
full[is.na(MiscFeature), MiscFeature := "none"]
```

Nominal, ordinal, and discrete variables are converted to `factor`,
continuous to `numeric`.

```{r}
for (j in c(nomVars, ordVars, discVars)) full[[j]] <- as.factor(full[[j]])
for (j in contVars) full[[j]] <- as.numeric(full[[j]])
```

**Ordering**

From *data_description* the order of ordinal variable classes can be derived.
These orderings are defined, then the `levels` of their `factors` are adjusted.

```{r}
lvlOrd <- list(
  LotShape = c("Reg", "IR1", "IR2", "IR3"),
  Utilities = c("AllPub", "NoSewr", "NoSeWa", "ELO"),
  LandSlope = c("Gtl", "Mod", "Sev"),
  BldgType = c("1Fam", "2FmCon", "Duplx", "TwnhsE", "TwnhsI"),
  OverallQual = 10:1,
  OverallCond = 10:1,
  ExterQual = c("Ex", "Gd", "TA", "Fa", "Po"),
  ExterCond = c("Ex", "Gd", "TA", "Fa", "Po"),
  BsmtQual = c("Ex", "Gd", "TA", "Fa", "Po", "noBasement"),
  BsmtCond = c("Ex", "Gd", "TA", "Fa", "Po", "noBasement"),
  BsmtExposure = c("Gd", "Av", "Mn", "No", "noBasement"),
  BsmtFinType1 = c("GLQ", "ALQ", "BLQ", "Rec", "LwQ", "Unf", "noBasement"),
  BsmtFinType2 = c("GLQ", "ALQ", "BLQ", "Rec", "LwQ", "Unf", "noBasement"),
  HeatingQC = c("Ex", "Gd", "TA", "Fa", "Po"),
  KitchenQual = c("Ex", "Gd", "TA", "Fa", "Po"),
  Functional = c("Typ", "Min1", "Min2", "Mod", "Maj1", "Maj2", "Sev", "Sal"),
  FireplaceQu = c("Ex", "Gd", "TA", "Fa", "Po", "noFireplace"),
  GarageFinish = c("Fin", "RFn", "Unf", "noGarage"),
  GarageQual = c("Ex", "Gd", "TA", "Fa", "Po", "noGarage"),
  GarageCond = c("Ex", "Gd", "TA", "Fa", "Po", "noGarage"),
  PoolQC = c("Ex", "Gd", "TA", "Fa", "noPool"),
  Fence = c("GdPrv", "MnPrv", "GdWo", "MnWw", "noFence"),
  MoSold = 1:12
)
for (j in names(lvlOrd)) {
  full[[j]] <- factor(full[[j]], levels = lvlOrd[[j]])
}
```

## Feature Dependencies

Some features' values heavily depend on the value of another feature.
For example the values of $BsmtUnfSF$, $TotalBsmtSF$ only make sense if
features $BsmtCond$, $BsmtExposure$, $BsmtQual$ are not $noBasement$.
This information is given in *data_description*.

$$
\begin{aligned}
BsmtUnfSF, TotalBsmtSF : BsmtCond, BsmtExposure, BsmtQual \ne noBasement\\
BsmtFinSF1 : BsmtFinType1 \ne noBasement\\
BsmtFinSF2 : BsmtFinType2 \ne noBasement \\
GarageYrBlt, GarageCars, GarageArea: GarageType, GarageFinish, GarageQual, 
GarageCond \ne noGarage \\
PoolArea : PoolQC \ne noPool \\
MiscVal : MiscFeature \ne none \\
MasVnrArea : MasVnrType \ne none
\end{aligned}
$$
For the garage and basement features there are several features which indicate
the absence of that feature.
If one of them indicates the absence of a feature, the others should too.
This can be tested with the difference between two sets.

```{r}
outersect <- function(x, y) sort(c(setdiff(x, y), setdiff(y, x)))
```

The differences between Ids of observations with *noBasement* are computed.

```{r}
outersect(which(full$BsmtCond == "noBasement"),
          which(full$BsmtExposure == "noBasement"))
outersect(which(full$BsmtCond == "noBasement"),
          which(full$BsmtQual == "noBasement"))
full[c(949, 1488, 2041, 2186, 2218, 2219, 2349, 2525), 
     .(Id, BsmtCond, BsmtExposure, BsmtQual, BsmtUnfSF, TotalBsmtSF)]
```

So here the *data_description* is not consistent.
The general condition is set to *noBasement* but there are values
for basement exposure, quality, and so on.
*noBasement* was translated from *NA*s.
Since for all the other basement related variables there are reasonable values
it might be that in these cases *NA* did not mean *noBasement*.

The same can be done for the garage related variables.

```{r}
outersect(which(full$GarageType == "noGarage"),
          which(full$GarageFinish == "noGarage"))
outersect(which(full$GarageType == "noGarage"),
          which(full$GarageQual == "noGarage"))
outersect(which(full$GarageType == "noGarage"),
          which(full$GarageCond == "noGarage"))
full[c(2127, 2577), 
     .(GarageYrBlt, GarageCars, GarageArea, GarageType, 
       GarageFinish, GarageQual, GarageCond)]
```

Here, most garage related features were missing.



**Mark Undefined Variables**

If e.g. *PoolQC* is *noPool*, then there should be a missing vlaue or
a zero for *PoolArea*.
If there actually is a value, this could either mean it is a mistake
or it means something special that is not clearly described.
To at least be consistent, all *NA*s will be set to 0.

```{r, results='hide'}
full[BsmtCond == "noBasement" & is.na(BsmtUnfSF), BsmtUnfSF := 0]
full[BsmtCond == "noBasement" & is.na(TotalBsmtSF), TotalBsmtSF := 0]
full[BsmtFinType1 == "noBasement" & is.na(BsmtFinSF1), BsmtFinSF1 := 0]
full[BsmtFinType2 == "noBasement" & is.na(BsmtFinSF2), BsmtFinSF2 := 0]
full[GarageType == "noGarage" & is.na(GarageYrBlt), GarageYrBlt := "0"]
full[GarageType == "noGarage" & is.na(GarageCars), GarageCars := "0"]
full[GarageType == "noGarage" & is.na(GarageArea), GarageArea := 0]
full[PoolQC == "noPool" & is.na(PoolArea), PoolArea := 0]
full[MiscFeature == "noFeature" & is.na(MiscVal), MiscVal := 0]
```






## Missing Data

$NA$ which are left must be true missing data.

```{r}
naFeats <- apply(full[, !"SalePrice"], 2, function(x) any(is.na(x)))
naObs <- apply(full[, !"SalePrice"], 1, function(x) any(is.na(x)))
```

**Features**

The total number of missing values per feature will be saved.
The features with at least one missing value are shown below.

```{r}
total <- apply(full[, naFeats, with = FALSE], 2, function(j) sum(is.na(j)))
df <- data.frame(
  missing = total / nrow(full),
  name = factor(names(total), levels = names(total)[order(total)]),
  type = featAttr[match(names(total), featAttr$name), type]
)
ggplot(df, aes(x = name, y = missing, fill = type)) +
  geom_bar(stat = "identity") +
  scale_y_continuous("proportion missing") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Features with Missing Values")
```

**Observations**

Below the number of missing values for observations that have missing values
is shown.
Many observations have 1 or 2, some have 3 or 4 missing values.

```{r}
total <- apply(full[naObs, ], 1, function(i) sum(is.na(i)))
df <- data.frame(
  y = total,
  name = full[naObs, Id]
)
highchart() %>%
    hc_chart(type = "column") %>%
    hc_title(text = "Observations with Missing Values") %>%
    hc_xAxis(title = list(text = "observations"), categories = df$name, 
             labels  = list(rotation = "-45")) %>%
    hc_yAxis(title = list(text = "missing"), align = "left") %>%
    hc_add_series(showInLegend = FALSE, data = list_parse(df[order(df$y), ]))
```













***



















# Feature Exploration


## Feature Distributions

Feature distributions are described for each type of variable separately.

### Continuous Variables

Values were log10 transformed to better capture their spread.
There are many zeros, partly because of the dependency to another variable.
These zeros are excluded during the transformation.

```{r, fig.width=10, fig.height=8, warning=FALSE}
dt <- melt(full, "Id", contVars)
ggplot(dt, aes(x = value)) +
  geom_histogram(bins = 30) +
  scale_x_continuous(trans = "log10") +
  facet_wrap(~ variable, scale = "free") +
  theme_bw()
```

Summaries for each distribution are computed.
There are many undefined variables which are set to zero.
This influences the description a lot.
Therefore, they are excluded.

```{r}
dt <- full[, contVars, with = FALSE]
df <- data.frame(
  Min = apply(dt, 2, function(x) min(x[x > 0], na.rm = TRUE)),
  Mean = apply(dt, 2, function(x) mean(x[x > 0], na.rm = TRUE)),
  Median = apply(dt, 2, function(x) median(x[x > 0], na.rm = TRUE)),
  Max = apply(dt, 2, function(x) max(x[x > 0], na.rm = TRUE))
)
```

Relative difference between mean and median (*relDelta*) are computed as an
indicator for skewness.
Also, relative cardinality and relative number of outliers are computed.
Outliers are defined as being differing at least $+/- 1.5 MAD$ from the median.

```{r}
df$relDelta <- abs(df$Mean - df$Median) / df$Mean
df$relCard <- apply(dt, 2, function(x) {
  length(unique(x[x > 0])) / length(x[x > 0])
})
df$relOut <- apply(dt, 2, function(x) {
  length(boxplot.stats(x[x > 0])$out) / length(x[x > 0])
})
```

*PoolArea* and *MiscVal* show a high proportion of outliers,
but this might be because there are not many total values for these variables.
Many supposably continuous variables have a quite low cardinality.

```{r}
DT::datatable(df, options = list(dom = "ftp", pageLength = 20)) %>% 
  formatStyle('relDelta', 
    color = styleInterval(c(0.2, 0.8), c('green', 'orange', 'red'))) %>%
  formatStyle('relCard', 
    color = styleInterval(c(0.2, 0.6), c('red', 'orange', 'green'))) %>%
  formatStyle('relOut', 
    color = styleInterval(c(0.05, 0.1), c('green', 'orange', 'red'))) %>%
  formatRound(c(2, 5:7), 3)
```


### Discrete Variables

Distributions are shown below.

```{r, fig.width=10, fig.height=10, warning=FALSE}
dt <- melt(full, "Id", discVars)
ggplot(dt, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free") +
  theme_bw()
```

As summary, the value range, and cardinality are computed.

```{r}
dt <- full[, discVars, with = FALSE]
df <- data.frame(
  Card = apply(dt, 2, function(x) length(unique(x[!is.na(x)]))),
  Min = apply(dt, 2, function(x) min(as.numeric(as.character(x)), na.rm = TRUE)),
  Max = apply(dt, 2, function(x) max(as.numeric(as.character(x)), na.rm = TRUE))
)
```

Apart from that, the modes are usually interesting.
For this a function has to be defined first.

```{r}
Mode <- function(x, fst = TRUE) {
  ux <- unique(x)
  tab <- tabulate(match(x, ux))
  modeIdx <- which.max(tab)
  if (fst) return(list(value = ux[modeIdx], freq = tab[modeIdx]))
  tab[modeIdx] <- 0
  modeIdx <- which.max(tab)
  return(list(value = ux[modeIdx], freq = tab[modeIdx]))
}
```

First and second modes are computed and the table is shown.
Modes that make up more than 50% of observations will be highlighted.

```{r}
df$Mode <- apply(dt, 2, function(x) Mode(x)$value)
df$ModeCount <- apply(dt, 2, function(x) Mode(x)$freq)
df$Mode2nd <- apply(dt, 2, function(x) Mode(x, fst = FALSE)$value)
df$ModeCount2nd <- apply(dt, 2, function(x) Mode(x, fst = FALSE)$freq)
DT::datatable(df, options = list(dom = "ftp", pageLength = 20)) %>% 
  formatStyle('ModeCount', 
    color = styleInterval(nrow(dt) / 2, c("black", 'orange')))
```

In *GarageYrBlt* the maximum value is 2207, which does not make sense.
It seems that this is a typo and the real value is 2007.
After checking that there are not more values above 2010
2207 is changed to 2007.

```{r, results='hide'}
full[GarageYrBlt == 2207, GarageYrBlt := "2007"]
```


### Categorical Variables

**Ordinal Variables**

```{r, fig.width=10, fig.height=10, warning=FALSE}
dt <- melt(full, "Id", ordVars)
ggplot(dt, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free") +
  theme_bw()
```

**Nominal Variables**

```{r, fig.width=10, fig.height=10, warning=FALSE}
dt <- melt(full, "Id", nomVars)
ggplot(dt, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free") +
  theme_bw()
```

**Summary**

Cardinality, 1st and 2nd modes are given as summary.
Modes that make up more than 50% of observations will be highlighted.

```{r}
dt <- full[, c(ordVars, nomVars), with = FALSE]
df <- data.frame(
  Card = apply(dt, 2, function(x) length(unique(x[!is.na(x)]))),
  Mode = apply(dt, 2, function(x) Mode(x)$value),
  ModeCount = apply(dt, 2, function(x) Mode(x)$freq),
  Mode2nd = apply(dt, 2, function(x) Mode(x, fst = FALSE)$value),
  ModeCount2nd = apply(dt, 2, function(x) Mode(x, fst = FALSE)$freq),
  Type = rep(c("ord", "nom"), c(length(ordVars), length(nomVars)))
)
DT::datatable(df, options = list(dom = "ftp", pageLength = 20)) %>% 
  formatStyle('Type', 
    color = styleEqual(unique(df$Type), c('orange', 'blue'))) %>% 
  formatStyle('ModeCount', 
    color = styleInterval(nrow(dt) / 2, c("black", 'orange')))
```





## Correlations between Features

### Continuous

```{r}
vars <- c(contVars, discVars)
```

The Pearson Correlation Coefficient is calculated for all feature pairs
where both features are continuous.
The results are shown as heatmap below.
Using euclidean distance of the correlation matrix as a metric 
rows and columns were hierarchically ordered (complete link).

```{r, fig.width=10, fig.height=10}
m <- cor(
  full[, contVars, with = FALSE],
  use = "complete.obs"
)
corrplot(m, method = "color", order = "hclust")
```

Some clusters are visible.
The same hierarchical clustering is shown as tree below.

```{r}
corrClust <- hclust(dist(m))
plot(corrClust)
```


### Categorical

Here, discrete variables are included.

```{r}
vars <- c(discVars, ordVars, nomVars)
```

For categorical features Cramer's V is calculated for all feature pairs
to indicate correlation.
The results are shown as heatmap below.

```{r, fig.width=10, fig.height=10, warning=FALSE}
m <- as.matrix(full[, vars, with = FALSE])
m <- m[complete.cases(m), ]
mCramer <- matrix(NA, ncol(m), ncol(m))
for (i in seq_len(ncol(m))) {
  mCramer[i, ] <- vapply(
    seq_len(ncol(m)), 
    function(j) assocstats(table(m[, i], m[, j]))$cramer, 
    numeric(1)
  )
}
```

Using euclidean distance of the correlation matrix as a metric 
rows and columns were hierarchically ordered (complete link).

```{r}
colnames(mCramer) <- colnames(m)
rownames(mCramer) <- colnames(m)
corrplot(mCramer, method = "color", order = "hclust")
```

The tree for the clustering:

```{r}
corrClust <- hclust(dist(mCramer))
plot(corrClust)
```











***














# Imputation

## Single Imputation

Single missing values in basement related features, where all other
basement related features did have a value,
are replaced by their mode.

```{r, results='hide'}
full[c(949, 1488, 2349), BsmtExposure := "No"]
full[c(2041, 2186, 2525), BsmtCond := "TA"]
full[c(2218, 2219), BsmtQual := "TA"]
```

For garage related features there are two cases with almost no
non-missing value for garage.
For these two cases the garage features are set to *noGarage*.

```{r, results='hide'}
full[c(2127, 2577), GarageYrBlt := "0"]
full[c(2127, 2577), GarageCars := "0"]
full[c(2127, 2577), GarageArea := 0]
full[c(2127, 2577), GarageType := "noGarage"]
```

Features with only very few missing values will be imputed using the 
mode for categories and the median for continuous variables.

```{r, results='hide'}
full[is.na(Exterior1st), Exterior1st := "VinylSd"]
full[is.na(Exterior2nd), Exterior2nd := "VinylSd"]
full[is.na(Electrical), Electrical := "SBrkr"]
full[is.na(KitchenQual), KitchenQual := "TA"]
full[is.na(SaleType), SaleType := "WD"]
full[is.na(Utilities), Utilities := "AllPub"]
full[is.na(Functional), Functional := "Typ"]
full[is.na(MSZoning), MSZoning := "RL"]
full[is.na(MasVnrArea), MasVnrArea := 202]
full[is.na(MasVnrType), MasVnrType := "none"]
```

There were two observations with *NA* values for basement bath related variables
in which all other basement related variables were set to *noBasement*.
These will be set to zreo.

```{r, results='hide'}
full[is.na(BsmtFullBath), BsmtFullBath := "0"]
full[is.na(BsmtHalfBath), BsmtHalfBath := "0"]
```

## Multiple Imputation

*BldgType* and *LotFrontage* have around 10-15% missing values.
These will be imputed by Gibbs sampling of the posterior distribution.
Variables with missing values are repeatedly predicted given a set of predictor
variables.
For *LotFrontage* predictive mean matching, 
and for *BldgType* polytomous regression will be used.
As predictors all continuous variables and the the 3 categorical variables 
which are most correlated to *BldgType* are included.
Several datasets are created by this method but only one will be used.

```{r, results='hide'}
vars <- c(contVars, "BldgType", "MSSubClass", "BedroomAbvGr", "TotRmsAbvGrd")
dt <- full[, vars, with = FALSE]
imp <- mice(dt, seed = 42)
```

**LotFrontage**

The distribution of *LotFrontage* after imputation is compared to the 
original distribution.

```{r, warning=FALSE}
df <- data.frame(
  LotFrontage = c(dt$LotFrontage, complete(imp)$LotFrontage),
  Distro = rep(c("Original", "Imputed"), each = nrow(dt))
)
ggplot(df, aes(x = LotFrontage, color = Distro)) +
  geom_density() +
  theme_bw()
```

Both distributions look very similar.
Furthermore, its correlation with the 4 most correlated continuous variables
is compared.

```{r}
df <- data.frame(
  Value = c(dt$LotArea, dt$MasVnrArea, dt$GrLivArea, dt$GarageArea),
  Variable = rep(c("LotArea", "MasVnrArea", "GrLivArea", "GarageArea"), 
                 each = nrow(dt)),
  LotFrontage = rep(complete(imp)$LotFrontage, each = 4),
  isImputed = factor(is.na(dt$LotFrontage), levels = c(TRUE, FALSE))
)
ggplot(df, aes(x = LotFrontage, y = Value, color = isImputed))  +
  geom_point(size = .7) +
  facet_wrap( ~ Variable, scales = "free") +
  theme_bw()
```

Imputed values seem to resemble the shape of the original data.

**BldgType**

Imputed and original distributions are compared.

```{r, warning=FALSE, warning=FALSE}
df <- data.frame(
  BldgType = c(as.character(dt$BldgType), as.character(complete(imp)$BldgType)),
  Distro = rep(c("Original", "Imputed"), each = nrow(dt))
)
ggplot(df, aes(x = BldgType, fill = Distro)) +
  geom_bar(position = "dodge") +
  scale_x_discrete() +
  theme_bw()
```

Next correlation to the closest correlated categorical feature *MSSubClass*
before and after imputation is checked.

```{r}
df <- data.frame(
  BldgType = c(as.character(dt$BldgType), as.character(complete(imp)$BldgType)),
  MSSubClass = rep(dt$MSSubClass, 2),
  Distro = rep(c("Original", "Imputed"), each = nrow(dt))
)
ggplot(df, aes(x = MSSubClass, fill = BldgType)) +
  geom_bar(position = "dodge") +
  facet_grid(Distro ~ .) +
  theme_bw()
```

**Both**

Finally, the correlation of both imputed variables is checked.

```{r}
df <- data.frame(
  LotFrontage = rep(c(dt$LotFrontage, complete(imp)$LotFrontage), each = 2),
  BldgType = rep(c(as.character(dt$BldgType), 
                   as.character(complete(imp)$BldgType)), each = 2),
  Distro = rep(c("Original", "Imputed"), each = 2*nrow(dt))
)
ggplot(df, aes(x = LotFrontage, color = BldgType)) +
  geom_density() +
  facet_grid(Distro ~ .) +
  theme_bw()
```

The distribution of $BldgType = TwnhsE$ changed visibly.
However, with less than 10% this value is quite underrepresented
so its distribution estimate is less stable.
All in all the imputed values seem to be useful










***



