---
title: "Housing Cri.. Prices"
subtitle: "Model MARS"
author: "Marc Schwering"
output:
  html_document:
    number_sections: true
    toc: true
    toc_depth: 2
    fig_width: 10
    fig_height: 7
    theme: cosmo
---









# Pre-Processing

The following steps prepare the dataset for the modeling part.
The actual feature analysis was done in another kernel.

**Loading Data**

```{r, message=FALSE, warning=FALSE, results='hide'}
library(data.table) # database-like tables
library(ggplot2) # general plots
library(vcd) # stats for categories
library(mice) # imputation
library(earth) # MARS
library(crossval) # doing CV
test <- data.table(read.csv("../input/test.csv"))
train <- data.table(read.csv("../input/train.csv"))
test$SalePrice <- NaN
full <- rbind(train, test)
full$label <- rep(c("train", "test"), c(nrow(train), nrow(test)))
```

**Feature Attributes**

```{r, results='hide'}
# types
nomVars <- c("MSZoning", "Street", "Alley", "LandContour", "LotConfig",
             "Neighborhood", "Condition1", "Condition2", "HouseStyle",
             "RoofStyle", "RoofMatl", "Exterior1st", "Exterior2nd",
             "MasVnrType", "Foundation", "Heating", "CentralAir", "Electrical",
             "GarageType", "PavedDrive", "MiscFeature", "SaleType",
             "SaleCondition")
ordVars <- c("MSSubClass", "LotShape", "Utilities", "LandSlope", "BldgType",
             "OverallQual", "OverallCond", "ExterQual", "ExterCond", "BsmtQual",
             "BsmtCond", "BsmtExposure", "BsmtFinType1", "BsmtFinType2",
             "HeatingQC", "KitchenQual", "Functional", "FireplaceQu",
             "GarageFinish", "GarageQual", "GarageCond", "PoolQC", "Fence",
             "MoSold")
discVars <- c("YearBuilt", "YearRemodAdd", "BsmtFullBath", "BsmtHalfBath",
              "FullBath", "HalfBath", "BedroomAbvGr", "KitchenAbvGr",
              "TotRmsAbvGrd", "Fireplaces", "GarageYrBlt", "GarageCars",
              "YrSold")
contVars <- c("LotArea", "MasVnrArea", "BsmtFinSF1", "BsmtFinSF2", "BsmtUnfSF",
              "TotalBsmtSF", "X1stFlrSF", "X2ndFlrSF", "LowQualFinSF",
              "GrLivArea", "GarageArea", "WoodDeckSF", "OpenPorchSF",
              "EnclosedPorch", "X3SsnPorch", "ScreenPorch", "PoolArea",
              "MiscVal", "LotFrontage")
featAttr <- data.table(
  name = colnames(test)[-1],
  type = "nominal"
)
featAttr[name %in% ordVars, type := "ordinal"]
featAttr[name %in% discVars, type := "discrete"]
featAttr[name %in% contVars, type := "continuous"]

# special values
full[is.na(Alley), Alley := "noAccess"]
full[is.na(BsmtQual), BsmtQual := "noBasement"]
full[is.na(BsmtCond), BsmtCond := "noBasement"]
full[is.na(BsmtExposure), BsmtExposure := "noBasement"]
full[is.na(BsmtFinType1), BsmtFinType1 := "noBasement"]
full[is.na(BsmtFinType2), BsmtFinType2 := "noBasement"]
full[is.na(FireplaceQu), FireplaceQu := "noFireplace"]
full[is.na(GarageType), GarageType := "noGarage"]
full[is.na(GarageFinish), GarageFinish := "noGarage"]
full[is.na(GarageQual), GarageQual := "noGarage"]
full[is.na(GarageCond), GarageCond := "noGarage"]
full[is.na(PoolQC) , PoolQC := "noPool"]
full[is.na(Fence), Fence := "noFence"]
full[is.na(MiscFeature), MiscFeature := "none"]
full[GarageYrBlt == 2207, GarageYrBlt := 2007]

# variable types
for (j in c(nomVars, ordVars)) full[[j]] <- as.factor(full[[j]])
for (j in contVars) full[[j]] <- as.numeric(full[[j]])
for (j in discVars) full[[j]] <- as.integer(full[[j]])

# order levels
lvlOrd <- list(
  LotShape = c("Reg", "IR1", "IR2", "IR3"),
  Utilities = c("AllPub", "NoSewr", "NoSeWa", "ELO"),
  LandSlope = c("Gtl", "Mod", "Sev"),
  BldgType = c("1Fam", "2FmCon", "Duplx", "TwnhsE", "TwnhsI"),
  OverallQual = 10:1,
  OverallCond = 10:1,
  ExterQual = c("Ex", "Gd", "TA", "Fa", "Po"),
  ExterCond = c("Ex", "Gd", "TA", "Fa", "Po"),
  BsmtQual = c("Ex", "Gd", "TA", "Fa", "Po", "noBasement"),
  BsmtCond = c("Ex", "Gd", "TA", "Fa", "Po", "noBasement"),
  BsmtExposure = c("Gd", "Av", "Mn", "No", "noBasement"),
  BsmtFinType1 = c("GLQ", "ALQ", "BLQ", "Rec", "LwQ", "Unf", "noBasement"),
  BsmtFinType2 = c("GLQ", "ALQ", "BLQ", "Rec", "LwQ", "Unf", "noBasement"),
  HeatingQC = c("Ex", "Gd", "TA", "Fa", "Po"),
  KitchenQual = c("Ex", "Gd", "TA", "Fa", "Po"),
  Functional = c("Typ", "Min1", "Min2", "Mod", "Maj1", "Maj2", "Sev", "Sal"),
  FireplaceQu = c("Ex", "Gd", "TA", "Fa", "Po", "noFireplace"),
  GarageFinish = c("Fin", "RFn", "Unf", "noGarage"),
  GarageQual = c("Ex", "Gd", "TA", "Fa", "Po", "noGarage"),
  GarageCond = c("Ex", "Gd", "TA", "Fa", "Po", "noGarage"),
  PoolQC = c("Ex", "Gd", "TA", "Fa", "noPool"),
  Fence = c("GdPrv", "MnPrv", "GdWo", "MnWw", "noFence"),
  MoSold = 1:12
)
for (j in names(lvlOrd)) {
  full[[j]] <- factor(full[[j]], levels = lvlOrd[[j]])
}

# undefined values
full[BsmtCond == "noBasement" & is.na(BsmtUnfSF), BsmtUnfSF := 0]
full[BsmtCond == "noBasement" & is.na(TotalBsmtSF), TotalBsmtSF := 0]
full[BsmtFinType1 == "noBasement" & is.na(BsmtFinSF1), BsmtFinSF1 := 0]
full[BsmtFinType2 == "noBasement" & is.na(BsmtFinSF2), BsmtFinSF2 := 0]
full[GarageType == "noGarage" & is.na(GarageYrBlt), GarageYrBlt := 0]
full[GarageType == "noGarage" & is.na(GarageCars), GarageCars := 0]
full[GarageType == "noGarage" & is.na(GarageArea), GarageArea := 0]
full[PoolQC == "noPool" & is.na(PoolArea), PoolArea := 0]
full[MiscFeature == "noFeature" & is.na(MiscVal), MiscVal := 0]
```

**Missing Values**

```{r, results='hide'}
# single imputation
full[c(949, 1488, 2349), BsmtExposure := "No"]
full[c(2041, 2186, 2525), BsmtCond := "TA"]
full[c(2218, 2219), BsmtQual := "TA"]
full[c(2127, 2577), GarageYrBlt := 0]
full[c(2127, 2577), GarageCars := 0]
full[c(2127, 2577), GarageArea := 0]
full[c(2127, 2577), GarageType := "noGarage"]
full[is.na(Exterior1st), Exterior1st := "VinylSd"]
full[is.na(Exterior2nd), Exterior2nd := "VinylSd"]
full[is.na(Electrical), Electrical := "SBrkr"]
full[is.na(KitchenQual), KitchenQual := "TA"]
full[is.na(SaleType), SaleType := "WD"]
full[is.na(Utilities), Utilities := "AllPub"]
full[is.na(Functional), Functional := "Typ"]
full[is.na(MSZoning), MSZoning := "RL"]
full[is.na(MasVnrArea), MasVnrArea := 202]
full[is.na(MasVnrType), MasVnrType := "none"]
full[is.na(BsmtFullBath), BsmtFullBath := 0]
full[is.na(BsmtHalfBath), BsmtHalfBath := 0]

# transform
full$SalePrice <- log(full$SalePrice)
for (j in contVars) {
  full[[j]] <- log10(full[[j]] + 1)
}

# multiple imputation
vars <- c(contVars, discVars, "BldgType", "MSSubClass", "HouseStyle")
imp <- mice(full[, vars, with = FALSE], seed = 42)
full$BldgType <- complete(imp)$BldgType
full$LotFrontage <- complete(imp)$LotFrontage
```

**Important Features**

```{r}
impVars <- c("MSSubClass", "MSZoning", "LotFrontage", "LotArea", "LandContour", 
              "Neighborhood", "BldgType", "HouseStyle", "OverallQual", 
              "OverallCond", "YearBuilt", "YearRemodAdd", "Exterior1st", 
              "Exterior2nd", "MasVnrArea", "ExterQual", "Foundation", 
              "BsmtQual", "BsmtCond", "BsmtExposure", "BsmtFinType1", 
              "BsmtFinSF1", "BsmtUnfSF", "TotalBsmtSF", "HeatingQC", 
              "CentralAir", "X1stFlrSF", "X2ndFlrSF", "GrLivArea", 
              "BsmtFullBath", "FullBath", "HalfBath", "BedroomAbvGr", 
              "KitchenAbvGr", "KitchenQual", "TotRmsAbvGrd", "Fireplaces", 
              "FireplaceQu", "GarageType", "GarageYrBlt", "GarageFinish", 
              "GarageCars", "GarageArea", "GarageQual", "GarageCond", 
              "PavedDrive", "WoodDeckSF", "OpenPorchSF")
```

**Adjust Features**

```{r, results='hide'}
# binary features
full$noAccess <- full$Alley == "noAccess"
full$noBasement <- full$BsmtQual == "noBasement" | full$BsmtCond == "noBasement"
full$noFireplace <- full$FireplaceQu == "noFireplace"
full$noGarage <- full$GarageType == "noGarage" | full$GarageCond == "noGarage"
full$noPool <- full$PoolQC == "noPool"
full$noFence <- full$Fence == "noFence"
full$noMisc <- full$MiscFeature == "none"
binVars <- c("noAccess", "noBasement", "noFireplace", "noGarage", 
              "noPool", "noFence", "noMisc")

# ordinal to integer
for (j in ordVars) {
  full[[j]] <- as.integer(full[[j]])
}
```













***









# Modeling



## Preparation

The dataset is split back into training and test.

```{r}
train <- full[label == "train", !"label"]
test <- full[label == "test", !c("label", "SalePrice")]
X <- train[, !"SalePrice"]
Y <- train$SalePrice
```

Instead of using all features one could also use the set of features
labelled important by the Boruta analysis.
Then, the binary features could be added or removed from either set.

```{r}
Ximp <- X[, impVars, with = FALSE]
Xsmall <- X[, c(impVars, binVars), with = FALSE]
```

## Training


```{r}
library(xgboost)


# load data
data(agaricus.train, package='xgboost')
data(agaricus.test, package='xgboost')
train <- agaricus.train
test <- agaricus.test
# fit model
bst <- xgboost(data = train$data, label = train$label, max.depth = 1, eta = 1, nround = 3,
               nthread = 2, objective = "binary:logistic")
summary(bst)
# predict
pred <- predict(bst, test$data)

# plot trees
xgb.plot.tree(feature_names = agaricus.train$data@Dimnames[[2]], model = bst)
xgb.plot.multi.trees(model = bst, feature_names = agaricus.train$data@Dimnames[[2]], features.keep = 3)

# feature importance
importance_matrix <- xgb.importance(agaricus.train$data@Dimnames[[2]], model = bst)
xgb.plot.importance(importance_matrix)

# depth of leaves
xgb.plot.deepness(model = bst)

a































```



## Model Selection

Here, model selection can be directly done with the expected 
prediction error of the learning procedure.
This will be done using cross validation.
The root-mean-squared-error (RMSE) of the log of the target variable
is defined as error.

```{r}
L <- function(Y, Yh) sqrt(sum((Y - Yh)^2) / length(Y))
```

Inside the cross validation a prediction function will
train a MARS model on one part of the training set 
and then predict the other.
The RMSE for the prediction will be returned.

```{r}
predMARS <- function(trainX, trainY, testX, testY, ...) {
  fit <- earth(trainX, trainY, nk = 200, ...)
  pred <- predict(fit, newdata = testX)
  return(L(testY, pred[, 1]))
}
```

Splitting of the data and execution of the prediction function is performed
by `crossval`.
For each model it will repeat a 5-fold cross validation 10 times.
The resulting mean RMSE and its standard error are stored in a list.

```{r}
CV <- list()
CV$Add <- crossval(predMARS, X, Y, 
                   verbose = FALSE, K = 5, B = 10)
CV$Add_imp <- crossval(predMARS, Ximp, Y, 
                       verbose = FALSE, K = 5, B = 10)
CV$Add_small <- crossval(predMARS, Xsmall, Y, 
                         verbose = FALSE, K = 5, B = 10)
# CV$Int <- crossval(predMARS, X, Y, degree = 2, 
#                    verbose = FALSE, K = 5, B = 10)
# CV$Int_imp <- crossval(predMARS, Ximp, Y, degree = 2, 
#                        verbose = FALSE, K = 5, B = 10)
# CV$Int_small <- crossval(predMARS, Xsmall, Y, degree = 2, 
#                          verbose = FALSE, K = 5, B = 10)
```

Cases with interaction terms take to long to compute on kaggle, so they are
commented out.
Overall they performed worse than their additive counterparts.
Estimated RMSEs for all models are shown below.

```{r}
l <- lapply(CV, function(i) data.frame(RMSE = i$stat, se = i$stat.se))
res <- do.call(rbind, l)
res$Model <- rownames(res)
ggplot(res, aes(x = Model, y = RMSE)) +
  geom_pointrange(aes(ymax = RMSE + se, ymin = RMSE - se), color = "blue") +
  coord_flip() +
  ggtitle("CV-Estimated RMSE with Standard Error") +
  theme_classic()
```



## Final Model

*Add_imp* will be trained as final model.

```{r}
fit <- earth(Ximp, Y, nk = 200)
plot(fit)
summary(fit, decomp = "none")
```

The summary shows the terms in the order they were added during the
forward pass.
2 hinge functions for *OverallQual* and *GrLivArea* were added first.
These were also the most important features identified by the Boruta analysis.

Another way to estimate the variable importance is by looking at 
the number of terms in which they were used.
This is shown below in the column *nsubsets*.
The order of variables changed here compared to the summary.

```{r}
evimp(fit)[1:10, ]
```




















***





















# Prediction

Finally the test set can be predicted.

```{r}
Xtest <- test[, impVars, with = FALSE]
pred <- predict(fit, newdata = Xtest)[, 1]
```

The actual prices are shown below.

```{r}
prices <- exp(pred)
summary(prices)
ggplot(data.frame(Yh = prices), aes(x = Yh)) +
  geom_histogram(bins = 30) +
  scale_x_continuous(trans = "log10") +
  theme_bw()
```

That's it!
Thanks for reading through this whole thing.
If you have any suggestions, comments are welcome.

```{r}
df <- data.frame(
  Id = test$Id,
  SalePrice = prices
)
write.csv(df, file = "MARS_submission.csv", row.names = FALSE, quote = FALSE)
```




***
